---
title: "P8157 Homework 1 yz4184"
author: "Yunlin Zhou"
date: '2022-09-27'
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(lattice)
library(patchwork)
library(Hotelling)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```


```{r, import the data}
tlc = read.table("./TLC.dat") %>% 
  janitor::clean_names() %>% 
  rename(id = v1,
         assignment = v2,
         week0 = v3,
         week1 = v4,
         week4 = v5,
         week6 = v6)%>%
  mutate(assignment = as.factor(assignment))%>%
           as.data.frame()

zerbe2 = read.table("./ZERBE2.DAT") %>% 
  janitor::clean_names() %>% 
  rename(assignment = v1,
         id = v2,
         "0h" = v3,
         "0.5h" = v4,
         "1h" = v5,
         "1.5h" = v6,
         "2h" = v7,
         "3h" = v8)
```


# Part A

## 1

### (a) H0: ABC = D = 0

#### Matrix A

```{r}
matrix(c(1, -1), nrow = 1, ncol = 2, byrow = TRUE)
```

#### Matrix B

Matrix B is a 2*6 matrix consisting of group means at each time point.

#### Matrix C

```{r}
diag(6)
```


### (b) H0: ABC = D = 0

#### Matrix A

```{r}
matrix(c(1, -1), nrow = 1, ncol = 2, byrow = TRUE)
```

#### Matrix B

Matrix B is a 2*6 matrix consisting of group means at each time point.

#### Matrix C

```{r}
matrix(c(-1,0,0,0,0,
         1,-1,0,0,0,
         0,1,-1,0,0,
         0,0,1,-1,0,
         0,0,0,1,-1,
         0,0,0,0,1), nrow = 6, ncol = 5, byrow = TRUE)
```


### (c) H0: ABC = D = 0

#### Matrix A

```{r}
matrix(c(1, -1), nrow = 1, ncol = 2, byrow = TRUE)
```

#### Matrix B

Matrix B is a 2*2 matrix consisting of group means at 2 and 3 hours time point.

#### Matrix C

```{r}
matrix(c(-1, 1), nrow = 2, ncol = 1, byrow = TRUE)
```



# Part B

## 1

### EDA

```{r}
tlc_df = tlc %>% 
  pivot_longer(cols = week0:week6, 
               names_to = "week", 
               values_to = "blood_lead_levels_mg_dL") %>% 
  mutate(week = as.factor(week))
```


```{r}
tlc_p1 = ggplot(data = tlc_df, aes(x = week, 
             y = blood_lead_levels_mg_dL, 
             col = assignment, 
             group = id)) + geom_line()

tlc_p1
```


```{r}
tlc_p2 =  ggplot(data = tlc_df, aes(x = week, 
             y = blood_lead_levels_mg_dL, 
             col = assignment, 
             group = assignment)) + geom_smooth()

tlc_p2
```


```{r}
tlc_p3 =  ggplot(data = tlc_df, aes(x = week, 
             y = blood_lead_levels_mg_dL, 
             fill = assignment,
             alpha =0.3))+geom_boxplot()
  
tlc_p3 
```

### Hotelling's T2 test

```{r}
tlc_test1 = hotelling.test(.~assignment, data = tlc[,-1])
tlc_test1
```


```{r}
tlc2 = t(tlc[,-c(1:2)]) 
cmat = matrix(c(-1, 1, 0, 0, 0, -1, 1, 0, 0, 0, -1, 1), ncol = 4, byrow = TRUE)
tlc3 = cmat %*% tlc2
tlc3 = t(tlc3)
tlc3 = cbind(tlc[,2],tlc3) %>%
  data.frame()
tlc_test2 = hotelling.test(.~X1, data = tlc3)
tlc_test2
```


## 2

### Hotelling's T2 test

#### (a)

```{r}
z_test1 <- hotelling.test(.~assignment, data = zerbe2[,-2])
z_test1
```


#### (b)

```{r}
zerb_1 = t(zerbe2[,-c(1:2)])
cmat_z = matrix(c(-1,0,0,0,0,
         1,-1,0,0,0,
         0,1,-1,0,0,
         0,0,1,-1,0,
         0,0,0,1,-1,
         0,0,0,0,1), nrow = 6, ncol = 5, byrow = TRUE)%>%
  t()

zerb_2 = cmat_z %*% zerb_1
zerb_2 = t(zerb_2)
zerb_2 =cbind(zerbe2[,1], zerb_2)%>%
  data.frame()
test2 = hotelling.test(.~X1, data = zerb_2)
test2
```


#### (c)

```{r}
zerb_3 = t(zerbe2[, -c(1:6)])
cmat_z2 <- matrix(c(-1, 1), nrow = 2, ncol = 1, byrow = TRUE) %>% t()


zerb_4 = cmat_z2 %*% zerb_3
zerb_4 = t(zerb_4)
zerb_4 =cbind(zerbe2[,1], zerb_4)%>%
  data.frame()


test3 = hotelling.test(.~X1, data = zerb_4)
test3
```


### EDA


```{r}
z_df = zerbe2%>%
  pivot_longer(
    cols = "0h":"3h",
    names_to = "hours",
    values_to = "measurement"
  ) %>%
   mutate(hours = as.factor(hours),
          assignment = as.factor(assignment),
          id  = as.factor(id))%>%
  as.data.frame()
```


```{r}
z_p1 =  ggplot(data = z_df, aes(x = hours, 
             y = measurement, 
             col = assignment, 
             group = assignment)) + geom_smooth()

z_p1
```

We can conclude that the 2 groups are not parallel.

```{r}
z_p2 =  ggplot(data = z_df, aes(x = hours, 
             y = measurement, 
             fill = assignment,
             alpha =0.3))+geom_boxplot()
  
z_p2 
```

We can conclude that in control group, Plasma inorganic phosphate measurements are lower all the time. 

